---
title: "Generating a cohort"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Generating a cohort}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

This is a quick tutorial on using a cohort created using ATLAS and saving
cohort IDs and entry/exit in your user schema. This tutorial will use a cohort
created from the SynPuf synthetic dataset, using the cohort definition
"[RC] CVA Ischemic or Hemorrhagic with admission - synpuf". This cohort 
definition imported a phenotype for CVA from stroke ischemic or Hemorrhagic with 
admission from Phenotype library (https://data.ohdsi.org/PhenotypeLibrary/).
More details on generating cohorts can be found here: 
https://ohdsi.github.io/CohortGenerator/ and here: 
https://ohdsi.github.io/TheBookOfOhdsi/Cohorts.html. 

We'll need to install two new packages: the OHDSI Web Api and cohortGenerator

```{r setup, eval = FALSE}
remotes::install_github("OHDSI/ROhdsiWebApi")
remotes::install_github("OHDSI/CohortGenerator")
```

We'll use the same set up code as the intro vignette. 

```{r}
# ==============================================================================
# Packages =====================================================================
library(keyring)
library(DatabaseConnector)
library(tidyverse)
library(ohdsilab)
library(ROhdsiWebApi)
library(CohortGenerator)

# DB Connections ===============================================================
atlas_url = "https://atlas.roux-ohdsi-prod.aws.northeastern.edu/WebAPI"
cdm_schema = "omop_cdm_53_pmtx_202203"
synpuf_schema = "omop_cdm_synpuf_110k_531"
write_schema = paste0("work_", keyring::key_get("db_username"))
```

Instead of creating a database connection, we're just going to create a set of 
connection details. Some of the OHDSI R packages use this connectionDetails 
method instead of using a connection. 

```{r}
# Create connection details

Sys.setenv("DATABASECONNECTOR_JAR_FOLDER" = "insert path to jdbc driver here")

connectionDetails <- createConnectionDetails(dbms = "redshift",
                                             server = "ohdsi-lab-redshift-cluster-prod.clsyktjhufn7.us-east-1.redshift.amazonaws.com/ohdsi_lab",
                                             port = 5439,
                                             user = keyring::key_get("db_username"),
                                             password = keyring::key_get("db_password"))


# make it easier for some r functions to find the database
options(con.default.value = con)
options(schema.default.value = cdm_schema)
options(write_schema.default.value = write_schema)
```

We're also going to authorize the api for accessing ATLAS from R. This uses the
atlas_url information, and your username/password which you set using 
`keyring::key_set()` in a previous vignette ("Introduction to OHDSI Lab"). 

```{r}
ROhdsiWebApi::authorizeWebApi(atlas_url, 
                              authMethod = "db", 
                              webApiUsername = keyring::key_get("atlas_username"), 
                              webApiPassword = keyring::key_get("atlas_password"))
```


Ok now we need to do two things. First - pick which cohort to use to generate 
a data set. This is the number for the cohort that you (or someone else) creates
in ATLAS. Note that if you want to include multiple cohorts in the same table, 
you can provide a vector of cohort ID numbers.

```{r}
cohortId <- 1124
```

Then we're going to create a cohortDefinitionSet object using that ID and the 
baseUrl.

```{r}
cohortDefinitionSet <- ROhdsiWebApi::exportCohortDefinitionSet(baseUrl = atlas_url,
                                                               cohortIds = cohortId)

cohortDefinitionSet
```

We'll create a name for this cohort - this is the name that's used for the 
tables in your user schema.

```{r}
cohortTableNames <- getCohortTableNames(cohortTable = "synPuf_CVA")
cohortTableNames
```

We'll create empty tables to receive this data. We're using the 
connectionDetails object, the cohortTableName object, and the mySchema object 
from above. 

```{r, eval = FALSE}
# create empty tables in the {mySchema}.cohort table
createCohortTables(connectionDetails = connectionDetails,
                   cohortTableNames = cohortTableNames,
                   cohortDatabaseSchema = write_schema)
```

*This function doesn't actually currently work in Rmarkdown, see https://github.com/OHDSI/CohortGenerator/issues/97 for details. If you need to 
use it in RMarkdown supposedly you can wrap it in `purrr:quietly()`*

Finally, we'll add everyone from the database to our cohort. Notice that this 
code references the synpuf schema. If you were working with the pharmetrics 
data, you'd reference `cdm_schema` or `"omop_cdm_53_pmtx_202203"`.

```{r}
cohortsGenerated <- generateCohortSet(connectionDetails = connectionDetails,
                                      cdmDatabaseSchema = synpuf_schema,
                                      cohortDatabaseSchema = write_schema,
                                      cohortTableNames = cohortTableNames,
                                      cohortDefinitionSet = cohortDefinitionSet)
```

If you want to visually see this new table in your schema, you can connect to 
the database and use the **Connections** pane in the top right of RStudio. 

```{r}
con =  DatabaseConnector::connect(
  dbms = "redshift",
  server = "ohdsi-lab-redshift-cluster-prod.clsyktjhufn7.us-east-1.redshift.amazonaws.com/ohdsi_lab",
  port = 5439,
  user = keyring::key_get("db_username"),
  password = keyring::key_get("db_password"))
```

Then, you might use this table (for example) to reduce the full person table to 
include only people in your cohort.

```{r}
# pull data from the new cohort table
cohort <- tbl(con, inDatabaseSchema(write_schema, "synPuf_CVA")) |>
  select(person_id = subject_id, cohort_start_date, cohort_end_date)

# join to the person table to get information about demographics, calculate age at cohort entry
demographics <- cohort |> 
  omop_join("person", type = "inner", by = "person_id") |> 
  select(person_id, cohort_start_date, cohort_end_date, year_of_birth, gender = gender_source_value) |> 
  mutate(age_at_entry = year(cohort_start_date) - year_of_birth)

# if the above fails you may need to set the following options again first
options(con.default.value = con)
options(schema.default.value = cdm_schema)
options(write_schema.default.value = write_schema)

```

This code starts with a table made with CohortGenerator and uses it to limit the 
OMOP person table to just the people that we're interested in, as defined by our 
cohort, and then calculates an age at entry. 
