---
title: "Generating Covariates and Table 1"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Generating Covariates and Table 1}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This vignette summarizes the code that can be found here: http://ohdsi.github.io/FeatureExtraction/articles/UsingFeatureExtraction.html

I recommending working through the full vignette for those interested. This 
vignette will show how to apply the code to the ohdsi-lab database.

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

Start with the typical packages

```{r setup}
library(ohdsilab)
library(keyring)
library(DatabaseConnector)
library(ROhdsiWebApi)
library(CohortGenerator)
library(CohortDiagnostics)
library(FeatureExtraction)
```

Set the connection details

```{r}
atlas_url = "https://atlas.roux-ohdsi-prod.aws.northeastern.edu/WebAPI"
cdm_schema = "omop_cdm_53_pmtx_202203"
write_schema = paste0("work_", keyring::key_get("db_username"))

Sys.setenv("DATABASECONNECTOR_JAR_FOLDER" = "insert path to jdbc driver here")

connectionDetails <- createConnectionDetails(
	dbms = "redshift",
	server = "ohdsi-lab-redshift-cluster-prod.clsyktjhufn7.us-east-1.redshift.amazonaws.com/ohdsi_lab",
	port = 5439,
	user = keyring::key_get("db_username"),
	password = keyring::key_get("db_password"))
```

This code will generate a table 1 for an established cohort of individuals with 
concepts of stroke and aphasia. This table was generated using CohortGenerator 
and ATLAS, and saved as "RC_aphasia" in my own schema. In the 
`getDbCovariateData()` funtion, we need to provide information about the 
database connection (connectionDetails), the omop database data, the specific 
user schema to look in (`write_schema`). "subject_id" becomes the row for 
subject - we could also name the "person_id" for example. -1 says to calculate 
covariates for all cohorts in the table (there's only 1). 

The `aggregated` argument defaults to FALSE, so we'll get covariates at the 
person level here. We can aggregate them in the next step.

```{r}
# This sets the covarate settings as the defaults
covariateSettings <- createDefaultCovariateSettings()

# This pulls the covariate data from the cohort table
covariateData <- getDbCovariateData(
	connectionDetails = connectionDetails,
	cdmDatabaseSchema = cdm_schema,
	cohortDatabaseSchema = write_schema,
	cohortTable = "RC_aphasia",
	cohortIds = -1,
	rowIdField = "subject_id",
	covariateSettings = covariateSettings)

summary(covariateData)
```

Calculating covariates can be pretty time consuming, so it's good to save it 
once its generated. We can do this using 
```{r}
# save it like this. its an andromeda sql object, only nee dto do this once. 
saveCovariateData(covariateData, "covariates")

# we can then load later if needed it like this.
covariateData <- loadCovariateData("covariates")
```

According to the documentation, we should 'tidy up' the covariates. We can do 
this using the `tidyCovariateData()` function. 

*Normalize covariate values by dividing by the max and/or remove redundant covariates and/or remove infrequent covariates. For temporal covariates, redundancy is evaluated per time ID.*

Again, we'll save the result since it can take a while. 
```{r}
#this tidies up the covariates but I'm not sure what to use it for...
tidyCovariates <- tidyCovariateData(
	covariateData,
	minFraction = 0.001,
	normalize = FALSE,
	removeRedundancy = TRUE)

# aggregates summary statistics for table 1.
covariateData.agg <- aggregateCovariates(covariateData)
saveCovariateData(
	covariateData.agg,
	"covariates_agg")

# we can load it later like this:
covariateData.agg <- loadCovariateData("covariates_agg")
```

Finally, we can use the `createTable1()` function to put it all in a Table 1. I 
found that I prefer to have once column output. I haven't found a terribly 
satisfying way of getting the results out in a nicely formatted table, so for 
now I've just saved it as a .csv file and formatted in excel. 

```{r}
result <- createTable1(covariateData.agg)
print(
	result,
	row.names = FALSE,
	right = FALSE)
result <- createTable1(
	covariateData.agg,
	output = "one column")
write.csv(
	result,
	"covariates.csv",
	row.names = FALSE)
```




