---
title: "Generating Characterization Files"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Generating Characterization Files}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This vignette summarizes the code that can be found here: https://ohdsi.github.io/Characterization/ and illustrates how this code can be 
used within the OHDSI Lab against OHDSI Lab data.

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

Install one new package (Characterization) and load it along with the other
typical packages.

```{r setup}
renv::install("OHDSI/Characterization")

library(ohdsilab)
library(DatabaseConnector)
library(keyring)
library(ROhdsiWebApi)
library(CohortGenerator)
library(FeatureExtraction)
library(Characterization)
```

Set the connection details. I connect to the database here so I can set the 
option(con.default.value) setting in the next line.

```{r}
Sys.setenv("DATABASECONNECTOR_JAR_FOLDER" = "insert path to JDBC driver")

connectionDetails <- DatabaseConnector::createConnectionDetails(
	dbms = "redshift",
	server = "ohdsi-lab-redshift-cluster-prod.clsyktjhufn7.us-east-1.redshift.amazonaws.com/ohdsi_lab",
	port = 5439,
	user = keyring::key_get("db_username"),
	password = keyring::key_get("db_password"))

atlas_url = "https://atlas.roux-ohdsi-prod.aws.northeastern.edu/WebAPI"
synpuf_schema = "omop_cdm_synpuf_110k_531"
write_schema = paste0("work_", keyring::key_get("db_username"))

con =  DatabaseConnector::connect(connectionDetails)

options(con.default.value = con)
options(schema.default.value = synpuf_schema)
options(write_schema.default.value = write_schema)
```

Connect to ATLAS

```{r}
ROhdsiWebApi::authorizeWebApi(
	atlas_url,
	authMethod = "db",
	webApiUsername = keyring::key_get("atlas_username"),
	webApiPassword = keyring::key_get("atlas_password"))
```

Instead of setting one cohort Id, this time you're setting a target cohort Id 
and an outcome id as the purpose of the characterization package is to compare a
target cohort with an outcome cohort.

```{r}
targetId <- 4675
outcomeId <- 4681
```

Pull the cohorts definitions from ATLAS, and use them to populate tables in 
your write_schema (this code comes from the "Generating a cohort" vignette)

```{r}
cohortDefinitionSet <- ROhdsiWebApi::exportCohortDefinitionSet(
	baseUrl = atlas_url,
	cohortIds = c(targetId, outcomeId))


cohortTableNames <- CohortGenerator::getCohortTableNames(
	cohortTable = "cohort")


createCohortTables(
	connectionDetails = connectionDetails,
	cohortTableNames = cohortTableNames,
	cohortDatabaseSchema = write_schema)


cohortsGenerated <- CohortGenerator::generateCohortSet(
	connectionDetails = connectionDetails,
	cdmDatabaseSchema = synpuf_schema,
	cohortDatabaseSchema = write_schema,
	cohortTableNames = cohortTableNames,
	cohortDefinitionSet = cohortDefinitionSet)
```

Now, set the covariate settings. The createCovariateSettings() function can be
used to customize the output of characterization. Run ?createCovariateSettings 
to view all possible arguments. For now, we'll just use 
createDefaultCovariateSettings() as we don't need anything specific. The 
aggreegateCovariate Settings will be used as an argument when we set the 
characterization settings. Other argument options for characterization are 
timeToEventSettings and dechallengeRechallengeSettings.

```{r}
aggregateCovariateSettings <- Characterization::createAggregateCovariateSettings(
  targetIds = targetId,
  outcomeIds = outcomeId,
  riskWindowStart = 1,
  startAnchor = 'cohort start',
  riskWindowEnd = 365,
  endAnchor = 'cohort start',
  covariateSettings = FeatureExtraction::createDefaultCovariateSettings())
```

Set the characterization settings

```{r}
characterizationSettings <- Characterization::createCharacterizationSettings(
  aggregateCovariateSettings = aggregateCovariateSettings)
```

Run the characterization. This will generate a series of csv files that can 
later be analyzed using traditional or programmatic methods. You may want to 
create a folder before running the following code and set the path of 
outputDirectory and executionPath to that folder.

```{r}
Characterization::runCharacterizationAnalyses(
  connectionDetails = connectionDetails,
  cdmDatabaseSchema = synpuf_schema,
  targetDatabaseSchema = write_schema,
  targetTable = 'cohort',
  outcomeDatabaseSchema = write_schema,
  outcomeTable = 'cohort',
  characterizationSettings = characterizationSettings,   
  outputDirectory = "insert path to output folder",
  executionPath = "insert path to output folder",
  csvFilePrefix = 'insert custom prefix')
```

And that's it! The characterization files may take a while to generate depending
on the size of your cohorts and the amount of variables you include in your 
characterization settings.


