---
title: "Population Level Estimation using CohortMethod"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Population Level Estimation using CohortMethod}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

This tutorial will illustrate how to use the CohortMethod package created by the
OHDSI community and described here: https://ohdsi.github.io/CohortMethod/.
Researchers can use this package to calculate difference of risk of specific 
clinical outcomes between a target cohort and a comparator cohort.

The Pharmetrics dataset (containing real patient claims data) will be used in 
this tutorial to ensure that correlation between risk of outcome and presence of outcome is consistent (Synthetic data may not produce consistent relationships between covariates). Any person-level results should not be shared with 
individuals who are not active OHDSI Lab users.

We’ll need to install one new package: CohortMethod
```{r}
renv::install("OHDSI/CohortMethod")
```

We’ll generate a cohort from ATLAS as we did in vignette 3, except now we need 
three cohorts: a target cohort (i.e. the population we’re analyzing): 4862 (Metformin & insulin exposure with prior type 2 diabetes diagnosis), a comparator cohort 
(i.e. the population we're comparing with the target population): 4859 (Insulin exposure with prior type 2 diabetes diagnosis), and an outcome cohort (i.e. the outcome we’re calculating risk for): 4861 (Hypoglycemia).

```{r}
#load the necessary packages
library(CohortGenerator)
library(tidyverse)
library(keyring)
library(CohortMethod)
library(ROhdsiWebApi)
library(DatabaseConnector)
library(ohdsilab)

#set path to JDBC driver
Sys.setenv("DATABASECONNECTOR_JAR_FOLDER" = "INSERT PATH TO JDBC DRIVER")

#set database connection details
connectionDetails <- createConnectionDetails(
	dbms = "redshift",
	server = "ohdsi-lab-redshift-cluster-prod.clsyktjhufn7.us-east-1.redshift.amazonaws.com/ohdsi_lab",
	port = 5439,
	user = keyring::key_get("db_username"),
	password = keyring::key_get("db_password"))

#set atlas url and schema names
atlas_url = "https://atlas.roux-ohdsi-prod.aws.northeastern.edu/WebAPI"
cdm_schema = "omop_cdm_53_pmtx_202203"
write_schema = paste0("work_", keyring::key_get("db_username"))

#connect to the ohdsilab database and setting 
con =  DatabaseConnector::connect(connectionDetails)

#make it easier for some r functions to find the database and schemas
options(con.default.value = con)
options(schema.default.value = cdm_schema)
options(write_schema.default.value = write_schema)

#establish authorization to connect to ATLAS API
ROhdsiWebApi::authorizeWebApi(
  atlas_url,
  authMethod = "db",
  webApiUsername = keyring::key_get("atlas_username"),
  webApiPassword = keyring::key_get("atlas_password"))

#set the cohort IDs (target, comparator, and outcome)
cohortIds <- c(4859,4861,4862)

#extract the cohort definitions from ATLAS
cohortDefinitionSet <- ROhdsiWebApi::exportCohortDefinitionSet(
  baseUrl = atlas_url,
  cohortIds = cohortIds)

#set a nameing convention for the cohort tables
cohortTableNames <- getCohortTableNames(cohortTable = "T2D_cohortMethod")

#create empty tables in your personal schema using the naming convention
#designated in the last step.
createCohortTables(
  connectionDetails = connectionDetails,
  cohortTableNames = cohortTableNames,
  cohortDatabaseSchema = write_schema)

#generate your cohorts for the pmtx database.
cohortsGenerated <- generateCohortSet(
  connectionDetails = connectionDetails,
  cdmDatabaseSchema = cdm_schema,
  cohortDatabaseSchema = write_schema,
  cohortTableNames = cohortTableNames,
  cohortDefinitionSet = cohortDefinitionSet)
```

Now we're ready to start desiging the CohortMethod analysis. First, we create covariates using Feature Extraction as shown in vignette 6, making sure to
exclude the concept IDs of your study drugs (i.e. the drugs included in your
target and comparator cohorts) and those drugs' descendant concepts.
```{r}
covSettings <- createDefaultCovariateSettings(
  excludedCovariateConceptIds = c(30361,1503297,21600713,21500149,1992733),
  addDescendantsToExclude = TRUE
)
```

Extract the relevant data from the database. This may take a while.
```{r}
cohortMethodData <- getDbCohortMethodData(
  connectionDetails = connectionDetails,
  cdmDatabaseSchema = cdm_schema,
  targetId = 4862,  # ID for drugA cohort
  comparatorId = 4859,  # ID for drugB cohort
  outcomeIds = 4861,  # ID for outcome cohort
  exposureDatabaseSchema = write_schema,
  exposureTable = "T2D_cohortMethod",
  outcomeDatabaseSchema = write_schema,
  outcomeTable = "T2D_cohortMethod",
  covariateSettings = covSettings
)
```

Because the extract takes so long, I recommend saving it in case you need to
come back to the later steps during another session. Once you have saved the
data, reload it as the same variable.
```{r}
saveCohortMethodData(cohortMethodData, "cohortMethodCovariates.zip")
cohortMethodData <- loadCohortMethodData("cohortMethodCovariates.zip")
```

Create the study population with the appropriate risk windows. These can be
changed to match the needs of your study.
```{r}
studyPop <- createStudyPopulation(
  cohortMethodData = cohortMethodData,
  outcomeId = 4861,
  firstExposureOnly = FALSE,
  restrictToCommonPeriod = FALSE,
  washoutPeriod = 0,
  removeDuplicateSubjects = "keep all",
  removeSubjectsWithPriorOutcome = TRUE,
  minDaysAtRisk = 1,
  riskWindowStart = 0,
  startAnchor = "cohort start",
  riskWindowEnd = 30,
  endAnchor = "cohort end"
)
```

View attrition (how many subjects were excluded after each bit of inclusion
criteria?):
```{r}
getAttritionTable(studyPop)
```

Run the propensity score model (setting cross-validation to FALSE saves a lot of time, but this will still take a while):
```{r}
ps <- createPs(
  cohortMethodData = cohortMethodData,
  population = studyPop,
  maxCohortSizeForFitting = 150000,  # Caps the sample size for fitting
  prior = createPrior("laplace", useCrossValidation = FALSE, exclude = c(0)),
  control = createControl(
    noiseLevel = "quiet",
    tolerance = 1e-06  # Slightly looser convergence criteria
  )
)
```

Because the propensity score model took so long to run, save and reload it.
```{r}
saveRDS(ps, "propensity_scores.rds")
ps <- readRDS("propensity_scores.rds")
```

Evaluate the propensity score model through the following 3 tests:
1. Calculate AUC (area under the curve)
```{r}
computePsAuc(ps)
```

2. Plot propensity score distribution
```{r}
plotPs(ps, 
       scale = "preference",
       showCountsLabel = TRUE,
       showAucLabel = TRUE,
       showEquiposeLabel = TRUE
)
```

3. Check which variables are in the model
```{r}
psModel <- getPsModel(ps, cohortMethodData)
```

Next, you should adjust your study population for confounding (I use the 
stratifyByPs methodbut you could also use matchOnPs or trimByPsToEquipoise 
depending on your study needs). Then visualize the adjusted study population.
```{r}
stratifiedPop <- stratifyByPs(ps, numberOfStrata = 5)
plotPs(stratifiedPop, ps, scale = "preference")
```

The next four steps evaluate the balance of your covariates.
1. Calculate the covariate balance
```{r}
balance <- computeCovariateBalance(stratifiedPop, cohortMethodData)
```

2. Visualize the covariate balance using a scatter plot and a plot of the most
prevalent covariates.
```{r}
plotCovariateBalanceScatterPlot(balance, showCovariateCountLabel = TRUE, showMaxLabel = TRUE)

plotCovariateBalanceOfTopVariables(balance)
```

3. Create a table 1 of covariates comparing the study population before and
after adjustment.
```{r}
createCmTable1(balance)
```

4. See how different the adjusted study population is from the original study
population by checking generalizability.
```{r}
getGeneralizabilityTable(balance)
```

The next two steps provide and assessment of the follow-up and the statistical power of your model.
1. Calculate the minimum detectable relative risk:
```{r}
computeMdrr(
  population = stratifiedPop,
  modelType = "cox",
  alpha = 0.05,
  power = 0.8,
  twoSided = TRUE
)
```

2. Examine follow-up distribution:
```{r}
getFollowUpDistribution(population = stratifiedPop)
plotFollowUpDistribution(population = stratifiedPop)
```

Finally, you're ready to run the outcome model to calculate the difference of
risk between the two populations (target_ID and comparator_ID).
```{r}
#This is a simple outcome model (Cox regression with propensity adjustment).
#There are also options for more complex models built into the CohortMethod
#package.
outcomeModel <- fitOutcomeModel(
  population = stratifiedPop,
  modelType = "cox",
  stratified = TRUE  # For stratified populations
)
outcomeModel  #View results
```

Now that you have your CohortMethod results, you can visualize them in a variety
of ways.
1. Create Kaplan-Meier plot:
```{r}
plotKaplanMeier(stratifiedPop, includeZero = FALSE)
```

2. Create time-to-event plot:
```{r}
plotTimeToEvent(
  cohortMethodData = cohortMethodData,
  outcomeId = 4861,
  firstExposureOnly = FALSE,
  washoutPeriod = 0,
  removeDuplicateSubjects = "keep all",
  minDaysAtRisk = 1,
  riskWindowStart = 0,
  startAnchor = "cohort start",
  riskWindowEnd = 30,
  endAnchor = "cohort end"
)
```

That's it! Obviously, many of the packages settings can be adjusted to fit your
study needs. The above just provides an example of what a CohortMethod analysis
might look like when run against OHDSI Lab data.
