---
title: "Designing a network analysis using Strategus"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Designing a network analysis using Strategus}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

Certain studies may require data elements not included in OHDSI Lab datasets. In
these scenarios, you may want to run an OHDSI network study, engaging with
OMOP CDM databases hosted by other institutions. Your code can be sent to
collaborators at these other institutions for them to run against their database
and return the results to you. There are multiple ways to handle this exchange
of code/analysis processes. You can send R or SQL scripts directly or you can
take advantage of the Strategus package, which modularizes other HADES packages 
(e.g., Characterization or Cohort Diagnostics) and creates an analysis 
specification in JSON format to be sent to collaborator researchers. The benefits
of Strategus include reducing the number of files that need to be exchanged 
(i.e. without Strategus, multiple HADES packages may require multiple files)
and reducing the complexity of running these shared analyses (running a Strategus
analysis requires far fewer lines of R code than running its separate parts).

There are three stages to this tutorial:
1. Creating the Strategus analysis specification: this is the most complicated as
it requires knowledge of other HADES packages. This step is completed by you (the 
programming lead).
2. Executing the Strategus analysis specification: This step is completed by
your collaborators at other database-hosting institutions once they have received
your JSON analysis specification.
3. Viewing the results: This step is completed by you (the study lead) once you 
have received your collaborator's results.

Stage 1: Design the analysis
To ensure that Strategus runs properly, I recommend downloading the Strategus 
template renv.lock file, which will install all R dependencies including the 
OHDSI HADES libraries and Strategus.

```{r setup, eval = FALSE}
download.file(
  "https://raw.githubusercontent.com/ohdsi-studies/StrategusStudyRepoTemplate/main/renv.lock",
  "INSERT PATH TO R PROJECT DIRECTORY")
install.packages("renv")
renv::activate()
renv::restore()
```

For the purposes of this tutorial, our analysis specification will include code
to generate the same type 2 diabetes cohort used in other vignettes and run the 
cohort diagnostics package to return a basic characterization of that cohort.
Other HADES packages including patient-level-prediction and cohort-incidence can 
also be added as Strategus modules depending on the needs of your study.

We'll start with designing the cohort generator module. To begin, we need a
cohort from ATLAS, which we retrieve in the same way done in previous vignettes.

```{r}
#set the atlas web API url
atlas_url = "https://atlas.roux-ohdsi-prod.aws.northeastern.edu/WebAPI"

#connect to ATLAS
ROhdsiWebApi::authorizeWebApi(
	atlas_url,
	authMethod = "db",
	webApiUsername = keyring::key_get("atlas_username"),
	webApiPassword = keyring::key_get("atlas_password"))

#choose an ATLAS cohort definition id
cohortId <- 4675

#export the chosen ATLAS cohort definition from ATLAS
cohortDefinitionSet <- ROhdsiWebApi::exportCohortDefinitionSet(
	baseUrl = atlas_url,
	cohortIds = c(targetId))
```

Now, we need to create the cohort generator module
```{r}
#create empty cohort generator module (cgModule)
cgModule <- Strategus::CohortGeneratorModule$new()

# Create a cohort definition shared resource element from the cohort generator 
# module
cohortDefinitionSharedResource <- cgModule$createCohortSharedResourceSpecifications(
  cohortDefinitionSet = cohortDefinitionSet
)

# Create a module specification
cohortGeneratorModuleSpecifications <- cgModule$createModuleSpecifications(
  generateStats = TRUE
)
```

Next, we need to create the cohort diagnostics module
```{r}
#create empty cohort diagnostics module (cdModule)
cdModule <- Strategus::CohortDiagnosticsModule$new()

#specify cohort diagnostics settings
cohortDiagnosticsModuleSpecifications <- cdModule$createModuleSpecifications(
  runInclusionStatistics = TRUE,
  runIncludedSourceConcepts = TRUE,
  runOrphanConcepts = TRUE,
  runTimeSeries = FALSE,
  runVisitContext = TRUE,
  runBreakdownIndexEvents = TRUE,
  runIncidenceRate = TRUE,
  runCohortRelationship = TRUE,
  runTemporalCohortCharacterization = TRUE
)
```

Finally, we need to combine our modules into the analysis specification
JSON file. When run (see the following stage), this specification will generate 
the type 2 diabetes cohorts as well as diagnostics (simple characterization) files 
about that cohort.
```{r}
#combine modules into one analysis specification
analysisSpecifications <- Strategus::createEmptyAnalysisSpecificiations() |>
  Strategus::addSharedResources(cohortDefinitionSharedResource) |>
	Strategus::addModuleSpecifications(cohortGeneratorModuleSpecifications) |>
  Strategus::addModuleSpecifications(cohortDiagnosticsModuleSpecifications)

#save analysis specification to a JSON file
ParallelLogger::saveSettingsToJson(
	analysisSpecifications,
	file.path("INSERT PATH TO OUTPUT FOLDER", "analysis_settings.json"))

```

Stage 2: Execute the analysis
Once your collaborator receives the JSON analysis specification, they need to
provide information about their database so the analysis knows where to run. 
The code I provide below (pulled from the other vignettes) refers to our 
PharMetrics database, so it would need to be changed to match whatever database 
your collaborators have access to.
```{r}
#loading relevant packages
library(Strategus)
library(DatabaseConnector)
library(keyring)

Sys.setenv("DATABASECONNECTOR_JAR_FOLDER" = "INSERT PATH TO JDBC DRIVER")

connectionDetails <- createConnectionDetails(
	dbms = "redshift",
	server = "ohdsi-lab-redshift-cluster-prod.clsyktjhufn7.us-east-1.redshift.amazonaws.com/ohdsi_lab",
	port = 5439,
	user = keyring::key_get("db_username"),
	password = keyring::key_get("db_password"))

cdm_schema = "omop_cdm_53_pmtx_202203"
write_schema = paste0("work_", keyring::key_get("db_username"))
```

Next, we need detail our execution settings.
```{r}
#choose a folder for the results files to populate
outputFolder <- "INSERT PATH TO EXECUTION OUTPUT FOLDER"

#set schemas, tables, and directories for outputs
executionSettings <- createCdmExecutionSettings(
  workDatabaseSchema = write_schema,
  cdmDatabaseSchema = cdm_schema,
  cohortTableNames = CohortGenerator::getCohortTableNames(),
  workFolder = file.path(outputFolder, "work_folder"),
  resultsFolder = file.path(outputFolder, "results_folder"),
  minCellCount = 5
)

#save the execution settings to JSON file. This file can be used for any analysis
#being run against the database described above
ParallelLogger::saveSettingsToJson(
  object = executionSettings,
  file.path(outputFolder, "execution_settings.json")
)
```

Finally, you (or your collaborator) are ready to execute the study. This will
output results files to the outputFolder indicated in the last step.
```{r}
#load the analysis specification JSON file from stage 1
analysisSpecifications <- ParallelLogger::loadSettingsFromJson(
  fileName = "INSERT PATH TO ANALYSIS SETTINGS JSON FILE")

#load the execution settings JSON file from the last step
executionSettings <- ParallelLogger::loadSettingsFromJson(
  fileName = "INSERT PATH TO EXECUTION SETTINGS JSON FILE")


#execute the analysis
Strategus::execute(
  connectionDetails = connectionDetails,
  analysisSpecifications = analysisSpecifications,
  executionSettings = executionSettings
)
```

Stage 3: View the results
Once the analysis has been executed, your collaborator can either run the 
following code to view the results themselves, or they can send you the results 
files, for you to run the following code and view the results. To view the 
results, you will need write access to a schema called "study_results" within
a postgres database. For instructions on setting up a local postgres database,
watch [this video](https://www.youtube.com/watch?v=oWsAYx2R9RI).

First, set the connection to the postgres database. This is where the results
will be written to for viewing. Download/install the postgres JDBC driver [here](https://jdbc.postgresql.org/).
```{r}
#You will need to replace the following details with your own postgres database
#details
resultsConnectionDetails <- DatabaseConnector::createConnectionDetails(
  dbms     = "postgresql", 
  server   = "localhost/postgres", 
  user     = "postgres", 
  password = "ohdsi", 
  port     = 5432, 
  pathToDriver = "INSERT PATH TO JDBC DRIVER (NEEDS TO BE INSTALLED)"
)
```

Next, you need to create empty tables in the correct format for the results.
```{r}
resultsDataModelSettings <- Strategus::createResultsDataModelSettings(
  resultsDatabaseSchema = "study_results",
  resultsFolder = "INSERT PATH TO RESULTS FOLDER",
)

Strategus::createResultDataModel(
  analysisSpecifications = analysisSpecifications,
  resultsDataModelSettings = resultsDataModelSettings,
  resultsConnectionDetails = resultsConnectionDetails
)
```

Then you populate those empty tables with your results files.
```{r}
Strategus::uploadResults(
  analysisSpecifications = analysisSpecifications,
  resultsDataModelSettings = resultsDataModelSettings,
  resultsConnectionDetails = resultsConnectionDetails
)
```

From here, you can query your results using postgreSQL. Alternatively, you can 
view them in an interactive web application using the RShiny package.
```{r}
#load the shiny packages
library(ShinyAppBuilder)
library(OhdsiShinyModules)

# specify the modules used in your analysis specification
shinyConfig <- initializeModuleConfig() |>
  addModuleConfig(
    createDefaultAboutConfig()
  ) |>
  addModuleConfig(
    createDefaultDatasourcesConfig()
  ) |>
  addModuleConfig(
    createDefaultCohortGeneratorConfig()
  ) |>
  addModuleConfig(
    createDefaultCohortDiagnosticsConfig()
  )

# now create the shiny app and view the results
ShinyAppBuilder::createShinyApp(
  config = shinyConfig,
  connectionDetails = resultsConnectionDetails,
  resultDatabaseSettings = createDefaultResultDatabaseSettings(schema = "study_results"),
  title = "INSERT TITLE OF STUDY",
  studyDescription = "INSERT SHORT DESCRIPTION OF STUDY"
)
```

An interactive RShiny application should pop up, enabling you to click through 
the results of the different modules you ran. That's it! You designed an
analysis specification, executed it against a dataset, and viewed the results!
Again the purpose of this Strategus package is to make the exchange and 
execution of study code/methods between collaborators at different institutions 
easier and less prone to errors.
