---
title: "Population Level Estimation using SelfControlledCaseSeries"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Population Level Estimation using SelfControlledCaseSeries}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

This tutorial will illustrate how to use the SelfControlledCaseSeries package created by the
OHDSI community and described here: https://ohdsi.github.io/SelfControlledCaseSeries/.
Researchers can use this package to calculate of risk of specific 
clinical outcomes given specific exposures by looking at exposed and unexposed
periods within individual person records. This estimation method differs from the 
estimation method used in the CohortMethod package because individual persons
serve as their own control in SelfControlledCaseSeries rather than relying on a
comparator population as in CohortMethod. Among other benefits, this method
enables perfect matching by eliminating genetic, lifestyle, and disease severity
differences between cases and controls.

The Pharmetrics dataset (containing real patient claims data) will be used in 
this tutorial to ensure that correlation between risk of outcome and presence of 
outcome is consistent (Synthetic data may not produce consistent relationships 
between covariates). Any person-level results should not be shared with 
individuals who are not active OHDSI Lab users.

For this tutorial, we'll be running an example estimation study on a cohort of
persons exposed to the antiepileptic drug clonazepam. We will be 
using SelfControlledCaseSeries to calculate risk of hypoventilation for persons 
within our exposure cohort. Because hypoventilation is a known side effect of
clonazepam exposure, we should expect to see increased risk in our estimation
model.

We’ll need to install one new package: SelfControlledCaseSeries
```{r}
renv::install("OHDSI/SelfControlledCaseSeries")
```

We’ll generate a cohort from ATLAS as we did in vignette 3 for the outcome 
cohort (i.e. the outcome we’re calculating risk for): 4896 (Hypoventilation).

```{r}
#load the necessary packages
library(DatabaseConnector)
library(keyring)
library(ROhdsiWebApi)
library(CohortGenerator)
library(SelfControlledCaseSeries)

#set path to JDBC driver
Sys.setenv("DATABASECONNECTOR_JAR_FOLDER" = "INSERT PATH TO JDBC DRIVER")

#set database connection details
connectionDetails <- createConnectionDetails(
	dbms = "redshift",
	server = "ohdsi-lab-redshift-cluster-prod.clsyktjhufn7.us-east-1.redshift.amazonaws.com/ohdsi_lab",
	port = 5439,
	user = keyring::key_get("db_username"),
	password = keyring::key_get("db_password"))

#set atlas url and schema names
atlas_url = "https://atlas.roux-ohdsi-prod.aws.northeastern.edu/WebAPI"
cdm_schema = "omop_cdm_53_pmtx_202203"
write_schema = paste0("work_", keyring::key_get("db_username"))

#connect to the ohdsilab database and setting 
con =  DatabaseConnector::connect(connectionDetails)

#make it easier for some r functions to find the database and schemas
options(con.default.value = con)
options(schema.default.value = cdm_schema)
options(write_schema.default.value = write_schema)

#establish authorization to connect to ATLAS API
ROhdsiWebApi::authorizeWebApi(
  atlas_url,
  authMethod = "db",
  webApiUsername = keyring::key_get("atlas_username"),
  webApiPassword = keyring::key_get("atlas_password"))

#set the cohort ID (outcome cohort)
outcomeId <- 4896

#extract the cohort definitions from ATLAS
cohortDefinitionSet <- ROhdsiWebApi::exportCohortDefinitionSet(
  baseUrl = atlas_url,
  cohortIds = outcomeId)

#set a naming convention for the cohort tables
cohortTableNames <- getCohortTableNames(cohortTable = "pmtxHypoventilation")

#create empty tables in your personal schema using the naming convention
#designated in the last step.
createCohortTables(
  connectionDetails = connectionDetails,
  cohortTableNames = cohortTableNames,
  cohortDatabaseSchema = write_schema)

#generate your cohorts for the pmtx database.
cohortsGenerated <- generateCohortSet(
  connectionDetails = connectionDetails,
  cdmDatabaseSchema = cdm_schema,
  cohortDatabaseSchema = write_schema,
  cohortTableNames = cohortTableNames,
  cohortDefinitionSet = cohortDefinitionSet)
```

Before we can design our estimation study, we need to define the drug exposure
we're calculating risk for. In this example, we are using a single concept Id,
but alternatively we could generate a second "exposure" cohort for this drug.
```{r}
clonazepam <- 798874
```

Now we can tell SelfControlledCaseSeries to extract all necessary data for our 
analysis. Note, that we're pulling the outcome cohort from the write_schema
where it was generated, and the exposureId from the drug_era table within the
cdm_schema. The drug_era table defines the spans of time persons are assumed to 
be exposed to a particular active ingredient (in this case clonazepam).
```{r}
sccsData <- getDbSccsData(
  connectionDetails = connectionDetails,
  cdmDatabaseSchema = cdm_schema,
  outcomeDatabaseSchema = write_schema,
  outcomeTable = "pmtxEpilepsy",
  outcomeIds = outcomeId,
  exposureDatabaseSchema = cdm_schema,
  exposureTable = "drug_era",
  exposureIds = clonazepam)
```

Let's get a quick look at our new object sccsData using base R's summary() 
function.
```{r}
summary(sccsData)
```

Depending on how long the extration takes, it may be wise to save the sccsData
object in case you need to come back to the later steps during another session. 
Once you have saved the data, reload it as the same variable.
```{r}
saveSccsData(sccsData, "sccsData.zip")
sccsData <- loadSccsData("sccsData.zip")
```

Next, we need to create our study population. Remember, because 
SelfControlledCaseSeries uses individuals as their own control, this study
population is only drawing from one cohort (no comparator cohort).
```{r}
studyPop <- createStudyPopulation(sccsData = sccsData,
                                  outcomeId = outcomeId,
                                  firstOutcomeOnly = FALSE,
                                  naivePeriod = 180)
```

Let's plot our sccsData and studypop objects to get an initial view of how 
hypoventilation events change before and after clonazepam exposure 
(top pane of the resulting graph) and how many persons have continuous 
observation at different intervals before and after clonazepam exposure (lower 
pane of the resulting graph). It looks like there are slightly more 
hypoventilation events after clonazepam exposure, which makes sense according to 
existing clinical knowledge. Once we run the estimation model, we'll know for 
sure.
```{r}
plotExposureCentered(
	sccsData = sccsData,
	studyPopulation = studyPop)
```

Next, we need to define the covariate we're including in our model (all drug 
eras of clonazepam).
```{r}
covarClonazepam <- createEraCovariateSettings(
  label = "Exposure of interest",
  includeEraIds = clonazepam,
  start = 0,
  end = 0,
  endAnchor = "era end")
```

Now we take our covariate and use it to define our sccsIntervalData, which
represents the data in non-overlapping time intervals, with information on the 
outcome (hypoventilation) and covariate per interval.
```{r}
sccsIntervalData <- createSccsIntervalData(
  studyPop,
  sccsData,
  eraCovariateSettings = covarClonazepam)
```

Finally, we use the sccsIntervalData to fit our estimation model, and then we
run model to view the resulting risk calculations. According to the model, the 
estimated risk of hypoventilation for persons exposed to clonazepam 
(Incidence Rate Ratio) is 1.27 (with a 95% confidence interval between 1.08 and 
1.49), meaning persons are 27% (with a 95% confidence interval between 8% and 
49%) more likely to experience hypoventilation if they are exposed to 
clonazepam. This result is expected as it aligns with existing clinical 
knowledge about clonazepam.
```{r}
model <- fitSccsModel(sccsIntervalData)
model
```
