---
title: "Generating a Synthetic Dataset"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Generating a cohort with Capr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

This tutorial will illustrate how to use the Eunomia package
created by the OHDSI community and described here: 
https://ohdsi.github.io/PatientLevelPrediction/. Researchers can use this
package to create a synthetic OMOP CDM (Common Data Model) dataset for 
demonstration and testing purposes. This package may have limited usefulness 
within the OHDSI Lab due to thealready-included SynPUF synthetic dataset, but
the Eunomia dataset can also be loaded outside the OHDSI Lab on a local machine 
or other workspace. This can be useful if you prefer to develop and test 
analytical code on your personal computer.

We'll need to install one new package: Eunomia

```{r setup, eval = FALSE}
renv::install("OHDSI/Eunomia")

library(Eunomia)
library(DatabaseConnector)
```

Now we create our connection details. This step is far simpler than creating
connection details for the OHDSI Lab databases (PharMetrics and SynPUF).

```{r}
connectionDetails <- Eunomia::getEunomiaConnectionDetails()
```

We connect to the Eunomia database is the same way we connected to the OHDSI Lab
databases.

```{r}
con <- DatabaseConnector::connect(connectionDetails)
options(con.default.value = con)
```

Now we can use the Eunomia database in much the same ways as we have the OHDSI
Lab databases. Let's generate a cohort of patients with Otitis Media to prove it.
The only notable difference is the cohortDatabaseSchema and cdmDataBaseSchema
will be set to "main".

```{r}
# Load the necessary packages
library(ROhdsiWebApi)
library(CohortGenerator)

# Set ATLAS url
atlas_url = "https://atlas.roux-ohdsi-prod.aws.northeastern.edu/WebAPI"

# Connect to ATLAS
ROhdsiWebApi::authorizeWebApi(
	atlas_url,
	authMethod = "db",
	webApiUsername = keyring::key_get("atlas_username"),
	webApiPassword = keyring::key_get("atlas_password"))

# Identify the cohort definition
cohortId <- 4734

# Export the cohort definition from ATLAS
cohortDefinitionSet <- ROhdsiWebApi::exportCohortDefinitionSet(
	baseUrl = atlas_url,
	cohortIds = cohortId)

# Set the naming convention for the cohort tables
cohortTableNames <- getCohortTableNames(cohortTable = "cohort")

# Create the cohort tables
createCohortTables(
	connectionDetails = connectionDetails,
	cohortTableNames = cohortTableNames,
	cohortDatabaseSchema = "main")

# Populate the cohort tables with persons in the Eunomia dataset who meet the
# cohort definition
cohortsGenerated <- generateCohortSet(
	connectionDetails = connectionDetails,
	cdmDatabaseSchema = "main",
	cohortDatabaseSchema = "main",
	cohortTableNames = cohortTableNames,
	cohortDefinitionSet = cohortDefinitionSet)
```

From here, we can do anything we normally do with cohorts (see vignettes 3-9). 
The Eunomia package can be a quick way to test or demonstrate code intended to 
be run against an OMOP CDM dataset.
