---
title: "KRD - Creating a Demographic Table 1"
output: html_document
date: "2026-04-28"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

This tutorial outlines the process of creating a demographic table 1 from the Komodo
Research Dataset stored within the OHDSI Lab. For the following code to
function, you must be logged into an OHDSI Lab workspace. This tutorial assumes
you have already successfully connected to the OHDSI Lab's Amazon Redshift
database according to the instructions in the previous article "General - 
Connecting to the database" and generated a cohort according the the instructions in the previous article "KRD - Generating a Cohort".


Load the additional necessary packages.
```{r}
library(dplyr)
library(ohdsilab)
```

Generate your table 1. Most of the argument values (con, ".t2_cohort", 
write_schema, and komodo_schema) were created in the last article. A min_count
value of 20 will only return demographic covariates representing 20 or more 
patients.
```{r}
t2d_table1 <- k_table1(
  con = con,
  cohort_table = "t2d_cohort",
  write_schema = write_schema,
  komodo_schema = komodo_schema,
  min_count = 20
)
```

View your table 1.
```{r}
View(t2d_table1)
```

Download your table 1 as a .csv file.
```{r}
write.csv(t2d_table1, "t2d_demographics", row.names = FALSE)
```
