---
title: "Generating-a-cohort"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Generating-a-cohort}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

This will be a quick tutorial on using a cohort created using ATLAS and saving
cohort IDs and entry/exit in your user schema. This tutorial will use a cohort
created from the SynPuf synthetic dataset, using the cohort definition
"[RC] CVA Ischemic or Hemorrhagic with admission - synpuf". This cohort definition
imported a phenotype for CVA from stroke ischemic or Hemorrhagic with admission from 
Phenotype library (https://data.ohdsi.org/PhenotypeLibrary/). The steps for doing this
will be recorded elsewhere. 

We'll need to install two new packages: the OHDSI Web Api and cohortGenerator

```{r setup, eval = FALSE}
remotes::install_github("OHDSI/ROhdsiWebApi")
remotes::install_github("OHDSI/CohortGenerator")
```

We'll use the same set up code as the intro vignette. 

```{r}
# ==============================================================================
# Packages =====================================================================
library(keyring)
library(DatabaseConnector)
library(tidyverse)
library(ohdsilab)
library(ROhdsiWebApi)
library(CohortGenerator)

# DB Connections ===============================================================
base_url = "https://atlas.roux-ohdsi-prod.aws.northeastern.edu/WebAPI"
cdm_schema = "omop_cdm_53_pmtx_202203"
synpuf_schema = "omop_cdm_synpuf_110k_531"
my_schema = paste0("work_", keyring::key_get("lab_user"))
```

Instead of creating a database connection, we're just going to create a set of connection
details. Some of the OHDSI R packages use this connectionDetails method instead of
using a connection. 

```{r}
# Create connection details
connectionDetails <- createConnectionDetails(dbms = "redshift",
                                             server = "ohdsi-lab-redshift-cluster-prod.clsyktjhufn7.us-east-1.redshift.amazonaws.com/ohdsi_lab",
                                             port = 5439,
                                             user = keyring::key_get("lab_user"),
                                             password = keyring::key_get("lab_password"))

```

We're also going to authorize the api for accessing ATLAS from R. This uses the
base_url information, and your username/password which you set using `keyring::key_set()`
to know that its ok for you to access ATLAS. 

```{r}
ROhdsiWebApi::authorizeWebApi(base_url, 
                              authMethod = "db", 
                              webApiUsername = keyring::key_get("lab_user"), 
                              webApiPassword = keyring::key_get("lab_password"))
```


Ok now we need to do two things. First - pick which cohort to use to generate 
a data set. This is the number for the cohort that you (or someone else) creates
in ATLAS. 

```{r}
cohortId <- 1124
```

Then we're going to create a cohortDefinitionSet object using that ID and the baseUrl.

```{r}
cohortDefinitionSet <- ROhdsiWebApi::exportCohortDefinitionSet(baseUrl = base_url,
                                                               cohortIds = cohortId)

cohortDefinitionSet
```

We'll create a name for this cohort - this is the name that's used for the tables
in your user schema.

```{r}
cohortTableNames <- getCohortTableNames(cohortTable = "synPuf_CVA")
cohortTableNames
```

We'll create empty tables to receive this data. We're using the connectionDetails
object, the cohortTableName object, and the mySchema object from above. 

```{r, eval = FALSE}
# create empty tables in the {mySchema}.cohort table
createCohortTables(connectionDetails = connectionDetails,
                   cohortTableNames = cohortTableNames,
                   cohortDatabaseSchema = my_schema)
```

*This function doesn't actually currently work in Rmarkdown, see https://github.com/OHDSI/CohortGenerator/issues/97 
for details. If you need to use it in RMarkdown supposedly you can wrap it in 
`purrr:quietly()`*

Finally, we'll add everyone from the database to our cohort. Notice that this code
references the synpuf schema. If you were working with the pharmetrics data, you'd
reference `cdm_schema` or `"omop_cdm_53_pmtx_202203"`.

```{r}
cohortsGenerated <- generateCohortSet(connectionDetails = connectionDetails,
                                      cdmDatabaseSchema = synpuf_schema,
                                      cohortDatabaseSchema = my_schema,
                                      cohortTableNames = cohortTableNames,
                                      cohortDefinitionSet = cohortDefinitionSet)
```

