---
title: "intro-to-ohdsilab"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{intro-to-ohdsilab}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

# Introduction to {ohdsilab}

This package is intended to streamline working with the ohdsilab pharmetrics data
at the Roux Institute. It contains some functions and template code using 
existing OHDSI R packages like {DatabaseConnector} as well as standard R packages
like {dplyr}, {tidyr}, and {tibble}. This vignette will walk through some of the
basic tools in ohdsilab

## Installation

Installation only needs to be done once, unless you are updating the package

```{r, eval = FALSE}
install.packages("remotes")
remotes::install_github("roux-ohdsi/ohdsilab")
```

## Load the ohdsilab package

- it should also load {DatabaseConnector}, {keyring}, and other dependencies,
but you might want to explicitly load them for transparency.

```{r setup}
library(ohdsilab)
library(tidyverse)
library(DatabaseConnector)
library(keyring)
```

## Snippets

{ohdsilab} stores some basic [snippets](https://rstudio.github.io/rstudio-extensions/rstudio_snippets.html) for quick starting new scripts and projects. Follow the instructions in the link above to install
the snippets. They are located in the /snippets folder in the package. The easiest way to copy
them is copy text below the first line from the raw file on github: https://raw.githubusercontent.com/roux-ohdsi/ohdsilab/main/snippets/snippets. If you notice any
mistakes or have any suggestions for the snippets, feel free to create a pull request or 
start an issue on the github page. The snippet you'll find most useful to begin with is
probably ohdsi_new_script.

Current snippets:

Polished:

- ohdsi_new_script

In the words:

- ohdsi_new_project
- ohdsi_creds
- ohdsi_db
- ohdsi_db_variables
- ohdsi_db_connect
- ohdsi_api_auth
- ohdsi_api_cohort_definition

## Functions

These are in development and being updated frequently. 


- `dbi_collect()` is a different version of the `collect()` function that uses DBI instead of dbplyr. There are
occasional errors with the `collect()` function where not all rows of the query are returned because the query
times out. In these cases, use the `dbi_collect()` function. 
- `rb()` is a function to "ROLLBACK" in redshift SQL. This is necessary when there is a mistake in the dbplyr generated
SQL code. You can use `rb()` instead of disconnecting and reconnecting to the database.

- `omop_join()` is a shortcut for a join from an existing query (e.g., a dplyr pipe chunk) to an omop table without
having to reference the connection and schema. 

- `icd2omop()` and `icd2omop2()` map a dataframe or vector of icd 9 or 10 codes to OMOP concept codes
- `cpt2omop()` maps a vector of cpt codes to omop concept codes. It can be used for any standardized vocabulary,
not just CPT4. (ICD9/10 are not considered standardized)
- `str_insert()` and `icd_periods()` help to insert periods into icd codes when they come without. This
is typically a necessary step before mapping to OMOP codes. 

## Connecting to the database

If this is the first time connecting on your computer, you'll need to set your credientials before
it will work. The snippet ohdsi_creds will create that code for you. When you
run each line, it'll generate a pop up where you can enter your database/ohdsilab
username (something like usr999) and your password (which is in your email). After
setting the credentials, you shouldn't have to do it again very often. 

```{r}
# Credentials
keyring::key_set("lab_user", prompt = "Username for this workspace")
keyring::key_set("lab_password", prompt = "Password for this workspace")
```


The following code was generated using the ohdsi_new_script snippet. It creates
a template for the basic code needed to connect to the database. Now that you've 
set up your credentials, see if it'll run without errors. Note that you'll also 
need to be using RStudio within the Amazon workspace box or you need to use the 
global connect VPN. 

```{r}
# ==============================================================================
# Packages =====================================================================
library(keyring)
library(DatabaseConnector)
library(tidyverse)
library(ohdsilab)

# Credentials ==================================================================
usr = keyring::key_get("lab_user")
pw  = keyring::key_get("lab_password")

# DB Connections ===============================================================
base_url = "https://atlas.roux-ohdsi-prod.aws.northeastern.edu/WebAPI"
cdm_schema = "omop_cdm_53_pmtx_202203"
my_schema = paste0("work_", keyring::key_get("lab_user"))
# Create the connection
con =  DatabaseConnector::connect(
  dbms = "redshift",
  server = "ohdsi-lab-redshift-cluster-prod.clsyktjhufn7.us-east-1.redshift.amazonaws.com/ohdsi_lab",
  port = 5439,
  user = keyring::key_get("lab_user"),
  password = keyring::key_get("lab_password")
)
class(con)
# make it easier for some r functions to find the database
options(con.default.value = con)
options(schema.default.value = cdm_schema)

# End Setup ====================================================================
# ==============================================================================
```

```{r}

```

